opXRD: Open Experimental Powder X-ray Diffraction Database

### 2024-12-10 chapters/03_existing_datasets.tex
Updated histogram

To contextualize opXRD within the current environment of experimental powder diffraction data, the following list provides an overview of the largest crystal structure databases that offer access to experimental powder diffraction data. For an overview of these databases refer to Tab.$\eqref{tab:exp_databases}$ below. In this table, the column ``O.A.'' indicates whether or not the database is open access. The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.

\begin{table}[!htb]
\centering
\caption{Overview of experimental powder diffraction databases}
\label{tab:exp_databases}
\scalebox{0.85}{
\begin{adjustbox}{center}
\begin{tabular}{@{}llllllll@{}}
\toprule
\textbf{Name} & \textbf{No. patterns} & \textbf{O.A.} & \textbf{Comp.} & \textbf{Spg.} & \textbf{Lattice} & \textbf{Atom coords.} & \textbf{Year est.} \\
\midrule
Linus Pauling file                     & 21,700                    & \xmark          & \cmark        & \cmark        & \cmark        & \cmark          & 2002            \\
Powder Diffraction File \tablefootnote{The PDF lists the Material Platform for Data Science (MPDS) as a database source. Since the MPDS is hosted by the Pauling File project, there is likely significant overlap in the experimental patterns available in the PDF and the Linus Pauling File.} & 20,800 & \xmark & \cmark & \cmark & \cmark & \partialcheck{52} & 1941 \\
RRUFF                                  & 1290                      & \cmark          & \cmark        & \cmark        & \cmark        & \xmark          & 2006            \\
Crystallography Open Database          & 1052                      & \cmark          & \cmark        & \cmark        & \partialcheck{85}        & \partialcheck{85}          & 2003            \\
PowBase                                & 169                       & \cmark          & \cmark        & \xmark        & \xmark        & \xmark          & 1999            \\
\bottomrule
\end{tabular}
\end{adjustbox}}
\end{table}

\textbf{Linus Pauling File}:\cite{PaulingWeb} The Linus Pauling File is a largely commercial crystal structure database published and maintained by the Pauling File project \cite{villars2018}. It is currently distributed as Pearson Crystal data \cite{PearsonWeb} and the Materials Platform for Data Science (MPDS)\cite{MPDSWeb}. The database, first published in 2002, currently contains more than 534,000 crystal structures\cite{MPDSWeb} and 21,700 corresponding experimental powder diffraction patterns\cite{PearsonWeb}. 
This makes the Pauling file, to the best of our knowledge, the largest collection of experimental powder diffraction data available to researchers. As of November 2024, Pearson's crystal data is available to researchers through a purchase of a one-year license starting at a price point of \qty{2200}{\myeuro}. The MPDS is partially open, with the open MPDS data accessible through a web interface\cite{MPDSWeb}. Access to the full MPDS database can be purchased through a one-year license starting at \qty{2000}{\myeuro}. \\

@@

\textbf{RRUFF}: \cite{RRUFFWeb} The RRUFF Mineral Database, first published in 2006, provides detailed information on minerals, including their chemical compositions, crystallography, and spectroscopic data. \cite{lafuente2015} Managed by the University of Arizona, it was created to serve as a public repository for mineral identification and research. It contains \num{1290} powder diffraction patterns stemming from experiments each labeled with the lattice parameters and composition of the underlying structures. \\

\textbf{Crystallography Open Database:} \cite{CODWeb} The Crystallography Open Database (COD) is an open-access collection of crystal structures founded in 2003\cite{Graulis2009cod}. It currently provides over 500,000 crystal structures. Of these files, 1052 contains the experimental powder diffraction data that was used to determine the underlying crystal structures of the investigated samples. Hence, the experimental powder diffraction data contained in the COD is mostly labeled with the full crystal structure information. \\

>>>>

To contextualize opXRD within the current environment of experimental powder diffraction data, the below list provides an overview of the largest crystal structure databases that offer access to experimental powder diffraction data. For an overview of these databases refer to Tab.$\eqref{tab:exp_databases}$ below. In this table, the column ``O.A.'' indicates whether or not the database is open access. The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'' respectively.

\begin{table}[!htb]
\centering
\caption{Overview of experimental powder diffraction databases}
\label{tab:exp_databases}
\scalebox{0.84}{
\begin{adjustbox}{center}
\begin{tabular}{@{}llllllll@{}}
\toprule
\textbf{Name} & \textbf{No. patterns} & \textbf{O.A.} & \textbf{Comp.} & \textbf{Spg.} & \textbf{Lattice} & \textbf{Atom coords.} & \textbf{Year est.} \\
\midrule
Linus Pauling file                     & 21,700                    & \xmark          & \cmark        & \cmark        & \cmark        & \cmark          & 2002            \\
Powder Diffraction File \tablefootnote{The PDF lists the Material Platform for Data Science (MPDS) as a database source. Since the MPDS is hosted by the Pauling File project, there is likely significant overlap in the experimental patterns available in the PDF and the Linus Pauling File.} & 20,800 & \xmark & \cmark & \cmark & \cmark & \partialcheck{52} & 1941 \\
RRUFF                                  & 1290                      & \cmark          & \cmark        & \cmark        & \cmark        & \xmark          & 2006            \\
Crystallography Open Database          & 1052                      & \cmark          & \cmark        & \cmark        & \partialcheck{85}        & \partialcheck{85}          & 2003            \\
PowBase                                & 169                       & \cmark          & \cmark        & \xmark        & \xmark        & \xmark          & 1999            \\
\bottomrule
\end{tabular}
\end{adjustbox}}
\end{table}

\textbf{Linus Pauling File}:\cite{PaulingWeb} The Linus Pauling File is a largely commercial crystal structure database published and maintained by the Pauling File project \cite{villars2018}. It is currently distributed as Pearson Crystal data \cite{PearsonWeb} and the Materials Platform for Data Science (MPDS)\cite{MPDSWeb}. The database, first published in 2002, currently contains more than 534,000 crystal structures\cite{MPDSWeb} and 21,700 corresponding experimental powder diffraction patterns\cite{PearsonWeb}. 
This makes the Pauling file, to the best of our knowledge, the largest collection of experimental powder diffraction data available to researchers. As of November 2024, Pearson's crystal data is available to researchers through a purchase of a one-year license starting at a price point of \qty{2200}{\myeuro}\cite{PearsonBuy}. The MPDS is partially open, with the open MPDS data accessible through a web interface\cite{MPDSWeb}. API access to the full MPDS can be purchased through a one-year license starting at \qty{9500}{\myeuro}\cite{MPDSBuy}. When inquired about the availability of experimental powder diffraction data in the MPDS API, the Pauling File project stated that this data is not currently included. It could however be offered in the future at the request of customers. \\

@@

\textbf{RRUFF}: \cite{RRUFFWeb} The RRUFF Mineral Database, first published in 2006, provides detailed information on minerals, including their chemical compositions, crystallography, and spectroscopic data. \cite{lafuente2015} Managed by the University of Arizona, it was created to serve as a public repository for mineral identification and research. It contains \num{1290} powder diffraction patterns stemming from experiments each labeled with the lattice parameters and composition of the underlying structures. The RRUFF data is openly accessible on the official website \cite{RRUFFWeb}. \\

\textbf{Crystallography Open Database:} \cite{CODWeb} The Crystallography Open Database (COD) is an open-access collection of crystal structures founded in 2003\cite{Graulis2009cod}. It currently provides over 500,000 crystal structures. Of these files, 1052 contains the experimental powder diffraction data that was used to determine the underlying crystal structures of the investigated samples. Hence, the experimental powder diffraction data contained in the COD is mostly labeled with the full crystal structure information. The data is openly accessible in the form of .cif files on the official COD website\cite{CODWeb}. \\

### 2024-12-10 chapters/04_our_dataset.tex
Updated histogram

In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiments, some of them labeled with corresponding structural information. The following paragraphs detail the contribution of each research group. Most data was collected using \ce{Cu} radiaton sources which have a  $K_{\alpha1}$ wavelength of $\lambda=1.54056\text{\AA}$ and a $K_{\alpha2}$ wavelength of $\lambda=1.54439\text{\AA}$. We filtered the submitted datasets to exclude patterns with invalid features such as the pattern containing only one unique recorded angle, the pattern containing less than 50 recorded angles total, or all intensities being zero. \\

@@

%TODO: A paragraph on how data is correlated
The degree of variation between the patterns is different for each contribution. For example, the CRNS and the INT contributions each are collections that encompass many research projects over a large period of time and hence exhibit a high degree of variability between individual patterns. In contrast, the contributions by the USC and the LBNL contain many very similar patterns. The patterns in the USC datasets are similar because the underlying samples are all variations of \ce{CuNi} and \ce{CuAl} alloys collected on the same machine. The patterns submitted by the LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical processes. The following paragraphs describe each contribution individually. \\

@@

\begin{figure}[!htb]
    \centering
    \includegraphics[width=\linewidth]{figures/quantities_hist.png}
    \caption{Histograms detailing spacegroups of labeled data and number of entries, starting angle value and final angle value for all data}
    \label{fig:enter-label}
\end{figure}

@@

\subsubsection*{\textit{In situ} XRD (RT $\le$ T$_{annealing}$ $\le$ 850$\degree$C)}

The phase evolution of the Mn-Sb-O system varying with annealing temperature was measured using the \textit{in situ} XRD (Rigaku-SmartLab3kW XRD, Figure \ref{fig:stability} (a)). This tool operates with SmartLab Studio II software, which is able to measure the X-ray diffraction during the annealing process. This enables to show directly all the phase transitions when annealing in various atmospheres such as O$_2$, Ar and NH$_3$. This device can theoretically operate up to 1000$\degree$C in a 20\% O$_2$ atmosphere. For safety reasons, phase transitions are analyzed with the \textit{in situ} XRD tool up to 850$\degree$C in this work.

Most of the \textit{in situ} experiments were performed under an air-like 20\% O$_2$ and 80\% argon (Ar) environment is chosen (Ar flow: 50 sccm, O$_2$ flow: 10 sccm). When a 100\% Ar environment is fixed, an Ar flow of 60 sccm is input.

Concerning the geometry, the Bragg-Brentano (BB) mode is preferred because it is more adapted in the analysis of scarce phases (i.e., MnSb$_2$O$_6$ rutile). The step is 0.01$\degree$ and the scanning rate is 10$\degree$/min. The step gives information about the resolution and the scanning rate (or speed duration time) results in the quality of the signal-to-noise ratio.

Figure \ref{fig:temp} shows the temperature profile during the \textit{in situ} XRD measurements. The temperatures used to annalyze the crystal structure of the Mn-Sb-O system were chosen depending on the number of phase transitions appearing for a certain temperature range. As explained in details in chapter \ref{res}, few changes in the crystal structure appear between RT and 300$\degree$C and phase transitions appeared from 300$\degree$C until 850$\degree$C. No phase transition appeared when cooling down. Therefore, the crystal structure was measured every 100$\degree$C between RT and 300$\degree$C; every 50$\degree$C between 300$\degree$C and 850$\degree$C; and every 200$\degree$C when cooling down. The heating and cooling rates are fixed for all the experiments at 50$\degree$C/min and the holding time is fixed at 2 min.

>>>>

In collaboration with several other institutions we have collected a dataset of diffractograms stemming from experiments, some of them labeled with corresponding structural information. The following paragraphs detail the contribution of each research group. Most data was collected using \ce{Cu} radiaton sources which have a  $K_{\alpha1}$ wavelength of $\lambda=1.54056\text{\AA}$ and a $K_{\alpha2}$ wavelength of $\lambda=1.54439\text{\AA}$. We filtered the submitted datasets to exclude patterns with invalid features such as the pattern containing only one unique recorded angle or negative angles, the pattern containing less than 50 recorded angles total, or all intensities being zero. \\

@@

%TODO: A paragraph on how data is correlated
The degree of variation between the patterns is different for each contribution. For example, the CRNS and the INT contributions each are collections that encompass many research projects over a large period of time and hence exhibit a high degree of variability between individual patterns. In contrast, the contributions by the USC and the LBNL contain many very similar patterns. The patterns in the USC datasets are similar because the underlying samples are all variations of \ce{CuNi} and \ce{CuAl} alloys collected on the same machine. The patterns submitted by the LBNL are similar because they stem from in-situ recordings where several hundred or several thousand patterns were collected over time per sample while they were undergoing physical processes.

%The below histograms $\eqref{fig:histograms}$ show the distributions of spacegroups,

\begin{figure}[!htb]
    \centering
    \includegraphics[width=\linewidth]{figures/quantities_hist.png}
    \caption{Histograms detailing spacegroups of labeled data and number of entries, starting angle value and final angle value for all data}
    \label{fig:histograms}
\end{figure}

@@

The following paragraphs describe each contribution individually. \\

@@

In a second study, the phase evolution of the Mn-Sb-O system varying with annealing temperature was measured using the \textit{in situ}  Rigaku-SmartLab3kW diffractometer. This tool operates with SmartLab Studio II software, which can measure the X-ray diffraction during the annealing process. This enables directly showing all the phase transitions when annealing in various atmospheres such as \ce{O2}, \ce{Ar}, and \ce{NH3}. Phase transitions are analyzed with the \textit{in situ} XRD tool up to 850$\degree$C in this work. Most of the \textit{in situ} experiments were performed under an air-like 20\% \ce{O2} and 80\% argon (Ar) environment is chosen (Ar flow: 50 \si{sccm}, O$_2$ flow: 10 \si{sccm}). When a 100\% Ar environment is fixed, an Ar flow of 60 \si{sccm} is input. The Bragg-Brentano (BB) mode is preferred in terms of geometry because it is more adapted in the analysis of scarce phases such as  \ce{MnSb2O6} rutile. The angular step is 0.01$\degree$ and the scanning rate is 10$\degree$/min. The step gives information about the resolution and the scanning rate (or speed duration time) results in the quality of the signal-to-noise ratio. \\

The temperatures used to analyze the crystal structure of the Mn-Sb-O system were chosen depending on the number of phase transitions appearing for a certain temperature range. Few changes in the crystal structure appear between RT and 300$\degree$C and phase transitions appeared from 300$\degree$C until 850$\degree$C. No phase transition appeared when cooling down. Therefore, the crystal structure was measured every 100$\degree$C between RT and 300$\degree$C; every 50$\degree$C between 300$\degree$C and 850$\degree$C; and every 200$\degree$C when cooling down. The heating and cooling rates are fixed for all the experiments at 50$\degree$C/min and the holding time is fixed at 2 min.
